---
title: "GSS challange Less Data"
author: "Leah  Castro"
date: "Fall 2025"
output: html_document
---



```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The data says white people and other races are often more in favor of capital punishment while black people seem to be more often against it. 

## Now with LESS data (more variables give less data due to NA's) create the approate graphic displays to answer the question.





### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~race + cappun, data=mydata)
rowPerc(table2)
colPerc(table2)
```

The data states that 65.10% of white people, 42.63% black and 55.20 others were in favor of capital punishment. On the opposing, 34.90 were white, 57.37 were black and 44.80 was composed of other races.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`
```{r}
fisherXY(mydata$race,mydata$cappun,printAll = TRUE)
```

When I observed the inferential results, it gave me the P-value for comparison between White and Black, White and other, and Black and Other. For White and Black the P-value was so low 1.741e-05 which showed that they are statistically different, but for the White vs Others and the Black vs Others the P-values were 0.2565 and 8.06299 which shows us that they are not statistically different. However, in the master branch, when doing the fisher test it showed us that all the values were statistically different.

# Discussion and Conclusion

This process was similar to the one in the master branch, but there were fewer observations. This happened because we only found a statistical difference between Black and White participants, giving us just 821 observations. As a result, the p-values didn’t change much except between those two groups. Having fewer observations means there’s less data to analyze, which I also saw in the Fisher’s test results from the number2 branch. The biggest difference was between Black and White participants, with a very low p-value of about 0.000017. This shows that more White people supported capital punishment, while fewer Black people did. The p-value between White and Other groups was 0.2565, meaning there wasn’t much difference—both groups mostly supported capital punishment. The p-value between Black and Other groups was 0.06299, showing they were pretty similar. Both the graphs and data showed the same pattern: Black participants were mostly against capital punishment, while White and Other participants were more in favor of it.

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
